Gap Puzzles

### 2020-01-05 gap/article/llncs2e/article.tex
report

\subsection{Search Strategy}
\paragraph{}
For labeling we made an ordering function in \textbf{crescent\_order(+B)} predicate. This predicate orders the coordinates of the points
top-down and left-right, improving the labeling times for smaller boards.

\begin{figure}
    \begin{center}
        \includegraphics[scale=0.5]{images/crescent.png}
        \caption{Crescent ordering} \label{fig4}
    \end{center}
\end{figure}

\paragraph{}
For bigger boards, \textbf{crescent\_order(+B)} predicate is an expensive task, as it runs through the whole list of points to apply the restrictions.
For this reason, if the board is of order 15 or bigger, this predicate is not used.
\paragraph{}
We also added a random function that was given in the classes. This function allows \textbf{labeling} predicate to select the value in a random way.

@@

\section{Results}
\paragraph{}
Our program finds one solution for every order given. Given this, we tested it with all possible numbers from 1 to 30.
By doing this we discovered that the puzzle only has solutions with orders 9 and above. Here are two examples.

\begin{figure}
    \begin{center}
        \includegraphics[scale=0.35]{images/gap9.jpg}
        \caption{Solution Example} \label{fig3}
    \end{center}
    \begin{center}
        \includegraphics[scale=0.5]{images/gap11.jpg}
        \caption{Solution Example} \label{fig4}
    \end{center}
\end{figure}

\paragraph{}
After making multiple tries per order we ended up with this graph. It shows the amount of time taken by the program for
each order. From the graph you can see a pattern where the program takes exponentially more time to solve the puzzle. But then it
reaches a point where there is a drop in that time. This pattern repeats itself after its first instance and probably forever, as you increase order.

\begin{figure}
    \begin{center}
        \includegraphics[scale=0.6]{images/graph.png}
        \caption{Times graph} \label{fig5}
    \end{center}
\end{figure}

\paragraph{}
This irregular behaviour told us the solution was imperfect somewhere. With that knowledge, we felt incentivized to search new ways of labeling variables.
We then found we could add a random function to the labeling and once we implemented it, it showed incredible results. After more testing, we reached the conclusion
that the best way to search for solutions in smaller boards was to use \textbf{crescent\_order(+B)}, but to bigger boards it was a waste of time. With this in mind,
we added a condition where if the board was of order 14 or below, it would use \textbf{crescent\_order(+B)}.

@@

\section{Conclusions and Future Work}
\paragraph{}
We think we have achieved what was proposed to us at the start of this project. With this work we were able to learn how to think in a different way.
Since we use other paradigms of programming more frequently, when using prolog with restrictions we could not think of a solution to the problem as quickly as normal.
\paragraph{}
As previously said, the gap puzzle is a challenging one, but with the use of restrictions, we can find solutions in little time.
One significant thing we learned was that a change that might seem simple, as adding a random function, might increase efficiency by many times.
\paragraph{}
From the results we can conclude that our program is quite efficient and that is one of its advantages. But this is at the cost of certainty, since you only get random
solutions, maybe not the one you are looking for.

>>>>

\subsection{Search Strategy}
\paragraph{}
For labeling we made an ordering function in \textbf{crescent\_order(+List)} predicate. This predicate orders the coordinates of the points
down-top and right-left, improving the labeling times for smaller boards.

\begin{figure}
    \begin{center}
        \includegraphics[scale=0.5]{images/crescent.png}
        \caption{Point ordering} \label{fig4}
    \end{center}
\end{figure}

\paragraph{}
For bigger boards, \textbf{crescent\_order(+List)} predicate is an expensive task, as it runs through the whole list of points to apply the restrictions.
For this reason, if the board is of order 15 or bigger, this predicate is not used.
\paragraph{}
We also added a random function that was given in the classes. This function allows \textbf{labeling} predicate to select the value in a random way.

@@

\newpage

\section{Results}
\paragraph{}
Our program finds one solution for every order given. Given this, we tested it with all possible numbers from 1 to 30.
By doing this we discovered that the puzzle only has solutions with orders 9 and above. Here is one example of one board with order 11.

\begin{figure}
    \begin{center}
        \includegraphics[scale=0.35]{images/gap9.jpg}
        \caption{Solution Example} \label{fig3}
    \end{center}
\end{figure}

\paragraph{}
After making multiple tries per order we ended up with this graph. It shows the amount of time (in seconds) taken by the program for
each order. From the graph you can see a pattern where the program takes exponentially more time to solve the puzzle. But then it
reaches a point where there is a drop in that time. This pattern repeats itself after its first instance and probably forever, as you increase order.

\begin{figure}
    \begin{center}
        \includegraphics[scale=0.5]{images/graph.png}
        \caption{Times graph} \label{fig5}
    \end{center}
\end{figure}

\paragraph{}
This irregular behaviour told us the solution was imperfect somewhere. With that knowledge, we felt incentivized to search new ways of labeling variables.
We then found we could add a random function to the labeling and once we implemented it, it showed incredible results. After more testing, we reached the conclusion
that the best way to search for solutions in smaller boards was to use \textbf{crescent\_order(+List)}, but to bigger boards it was a waste of time. With this in mind,
we added a condition where if the board was of order 14 or below, it would use \textbf{crescent\_order(+List)}, but bigger boards it would not.

@@

\pagebreak
\section{Conclusions and Future Work}
\paragraph{}
We think we have achieved what was proposed to us at the start of this project. With this work we were able to learn how to think in a different way.
Since we use other paradigms of programming more frequently, when using prolog with restrictions we could not think of a solution to the problem as quickly as normal.
\paragraph{}
As previously said, the gap puzzle is a challenging one, but with the use of restrictions, we can find solutions in little time.
One significant thing we learned was that a change that might seem simple, as adding a random function, might increase efficiency by many times.
\paragraph{}
From the results we can conclude that our program is quite efficient and that is one of its advantages. But this is at the cost of certainty, since you only get random
solutions, maybe not the one you are looking for.
